```{r setup tidyverse primer, include = FALSE}
# Not positive if the book will compile without this, so I'm including it just
# to be safe.
library(tidyverse)
```

# A tidyverse primer

**Learning objectives:**

-   List the [**`tidyverse` design principles.**](#tidyverse-design-principles)
-   Explain what it means for the [`tidyverse` to be **designed for humans.**](#design-for-humans)
-   Describe how [**reusing existing data structures** can make functions easier to work with.](#reusing-existing-data-structures)
-   Explain what it means for a set of functions to be [**designed for the pipe.**](#designed-for-the-pipe)
-   Explain what it means for function to be [**designed for functional programming.**](#designed-for-functional-programming)
-   List some [**differences between a `tibble` and a base `data.frame`.**](#tibbles-vs.-data-frames)
-   Recognize how to [**use the `tidyverse` to read and wrangle data.**](#how-to-read-and-wrangle-data)

## Tidyverse design Principles {#tidyverse-design-principles}

The `tidyverse` has [four core design principles](https://design.tidyverse.org/unifying-principles.html):

1.  **Human centered:** Designed to promote human usability.
2.  **Consistent:** Learning how to use one function or package is as similar as another.
3.  **Composable:** Easily breakdown data challenges into smaller components with exploratory tools to find the best solution.
4.  **Inclusive:** Fostering a community of like-minded users (e.g. \#*rstats*)

## Design for Humans {#design-for-humans}

> "Programs must be written for people to read, and only incidentally for machines to execute."
>
> \- Hal Abelson

The `tidyverse` offers packages that are easily readable and understood by humans. It enables them to more easily achieve their programming goals.

Consider the `mtcars` dataset, which comprises fuel consumption and 10 aspects of autombile design and performance from 1973-1974. Previewing the first six rows of the data, we see:

```{r, echo = FALSE}
head(mtcars)
```

If we wanted to arrange these in ascending order based on the `mpg` and `gear` variables, how could we do this?

------------------------------------------------------------------------

The function `arrange()`, in the `dplyr` package of the `tidyverse`, takes a data frame and column names as such:

```{r, eval = FALSE}
arrange(.data = mtcars, gear, mpg)
```

------------------------------------------------------------------------

`arrange()`, and other tidyverse functions, use **names that are descriptive and explicit.** For general methods, there is a **focus on verbs,** as seen with the functions `pivot_longer()` and `pivot_wider()` in the `tidyr` package.

## Reusing existing data structures {#reusing-existing-data-structures}

> "You don't have to reinvent the wheel, just attach it to a new wagon."
>
> \- Mark McCormack

There are many different data types in R, such as matrices, lists, and data frames.[^a_tidyverse_primer-1] A typical function would take in data of some form, conduct an operation, and return the result.

[^a_tidyverse_primer-1]: For a more detailed discussion, see Hadley Wickham's [*Advanced R*](https://adv-r.hadley.nz/vectors-chap.html)

`tidyverse` functions most often operate on data structures called tibbles.

-   Traditional data frames can represent different data types in each column, and multiple values in each row.

-   Tibbles are a special data frame that have additional properties helpful for data analysis.

    -   Example: list-columns

------------------------------------------------------------------------

```{r tidyverse-resample}
boot_samp <- rsample::bootstraps(mtcars, times = 3)
boot_samp
class(boot_samp)
```

------------------------------------------------------------------------

The above example shows how to create bootstrap resamples of the data frame mtcars. It returns a tibble with a `splits` column that defines the resampled data sets.

This function **inherits data frame and tibble methods so other functions that operate on those data structures can be used.**

## Designed for the pipe {#designed-for-the-pipe}

The pipe operator, `%>%`, comes from the [`magrittr`](https://magrittr.tidyverse.org) package by [Stefan Milton Bache](http://stefanbache.dk), and is used to chain together a sequence of R functions. More specifically, **the pipe operator uses the value of the object on the left-hand side of the operator as the first argument on the operator's right-hand side.**

The pipe allows for highly readable code. Consider wanting to sort the `mtcars` dataset by the number of gears (`gear`) and then select the first ten rows. How would you do that?

------------------------------------------------------------------------

```{r no pipe arrange slice}
cars_arranged <- arrange(mtcars, gear)
cars_selected <- slice(cars_arranged, 1:10)

# more compactly
cars_selected <- slice(arrange(mtcars, gear), 1:10)
```

Using the pipe to substitute the left-hand side of the operator with the first argument on the right-hand side, we can get the same result as follows:

```{r pipe arrange slice}
cars_selected <- mtcars %>%
  arrange(gear) %>%
  slice(1:10)
```

------------------------------------------------------------------------

This approach with the pipe works because all the functions **return the same data structure (a tibble/data frame) which is the first argument of the next function.**

**Whenever possible, create functions that can be incorporated into a pipeline of operations.**

## Designed for Functional Programming {#designed-for-functional-programming}

Functional Programming is an approach to replace iterative (i.e. for) loops. Consider the case where you may want two times the square root of the `mpg` for each car in `mtcars`. You could do this with a for loop as follows:

```{r for loop sqrt }
n <- nrow(mtcars)
roots <- rep(NA_real_, n)
for (car in 1:n) {
  roots[car] <- 2 * sqrt(mtcars$mpg[car])
}
```

You could also write a function to do the computations. In functional programming, it's important that the function **does not have any side effects and the output only depends on the inputs.** For example, the function `my_sqrt()` takes in a car's mpg and a weight by which to multiply the square root.

```{r my-sqrt}

my_sqrt <- function(mpg, weight) {
  weight * sqrt(mpg)
}

```

Using the [`purrr`](http://purrr.tidyverse.org/) package, we can forgo the for loop and use the `map()` family of functions which use the basic syntax of `map(vector, function)`. Below, we are applying the `my_sqrt()` function, with a weight of 2, to the first three elements of `mtcars$mpg`. User supplied functions can be declared by prefacing it with `~` (pronounced "twiddle").

-   By default, `map()` returns a list.

    -   If you know the class of a function's output, you can use special suffixes. A character output, for example, would used by `map_chr()`, a double by `map_dbl()`, and a logical by `map_lgl()`.

```{r map sqrt}
map(
  .x = head(mtcars$mpg, 3),
  ~ my_sqrt(
    mpg = .x,
    weight = 2
  )
)
```

-   `map()` functions can be used with 2 inputs, by specifying `map2()`

    -   Requires arguments `.x` and `.y`

```{r map2 sqrt}
map2(
  .x = head(mtcars$mpg, 3),
  .y = c(1,2,3),
  ~ my_sqrt(
    mpg = .x,
    weight = .y
  )
)
```

## Tibbles vs. Data Frames {#tibbles-vs.-data-frames}

A `tibble` is a special type of data frame with some additional properties. Specifically:

-   **Tibbles work with column names that are not syntactically valid variable names**.

```{r tibble syntax}
data.frame(`this does not work` = 1:2,
           oops = 3:4)

tibble(`this does work, though` = 1:2,
       `woohoo!` = 3:4)
```

-   **Tibbles prevent partial matching of arguments** to avoid accidental errors

```{r accidental matching tibble}
df <- data.frame(partial = 1:5)
tbbl <- tibble(partial = 1:5)

df$part

tbbl$part
```

-   **Tibbles prevent dimension dropping**, so subsetting data into a single column will never return a vector.

```{r subsetting vector conversion}
df[, "partial"]

tbbl[, "partial"]
```

-   **Tibbles allow for list-columns**, which can be a powerful tool when working with the `purrr` package.

```{r tibble list columns}

template_list <- list(a = 1, b = 2, c = 3, d = 4, e = 5)

data.frame(col = 1:5, list_col = template_list)

tibble(col = 1:5, list_col = template_list)

```

## How to read and wrangle data {#how-to-read-and-wrangle-data}

The following example shows how to use the `tidyverse` to read in data (with the `readr` package) and easily manipulate it (using the `dplyr` and `lubridate` packages). We will walk through these steps during our meeting.

```{r read wrangle data, message = FALSE}
library(tidyverse)
library(lubridate)

url <- "http://bit.ly/raw-train-data-csv"

all_stations <- 
  # Step 1: Read in the data.
  readr::read_csv(url) %>% 
  # Step 2: filter columns and rename stationname
  dplyr::select(station = stationname, date, rides) %>% 
  # Step 3: Convert the character date field to a date encoding.
  # Also, put the data in units of 1K rides
  dplyr::mutate(date = lubridate::mdy(date), rides = rides / 1000) %>% 
  # Step 4: Summarize the multiple records using the maximum.
  dplyr::group_by(date, station) %>% 
  dplyr::summarize(rides = max(rides), .groups = "drop")
```

```{r preview all stations}
head(all_stations, 10)
```

> "This pipeline of operations illustrates why the tidyverse is popular. A series of data manipulations is used that have simple and easy to understand user interfaces; the series is bundled together in a streamlined and readable way. The focus is on how the user interacts with the software. This approach enables more people to learn R and achieve their analysis goals, and adopting these same principles for modeling in R has the same benefits."
>
> \- Max Kuhn and Julia Silge in *Tidy Modeling with R*
