# Inferential analysis

**Learning objectives:**

- Use `broom::tidy()` to interpret the results of hypothesis tests.
- Use the `{infer}` package to test hypotheses.
- Use `rsample::reg_intervals` to compute bootstrap confidence intervals.
- Use `dplyr::mutate()` and `purrr::map()` to analyze parameter importance.

## Dataset used for demonstrating inference

I will use TidyTuesday dataset on ultra trail running races.

The data comes from [Benjamin Nowak](https://twitter.com/BjnNowak) by way of [International Trail Running Association (ITRA)](https://itra.run/Races/FindRaceResults). Their original repo is available on [GitHub](https://github.com/BjnNowak/UltraTrailRunning).

```{r 21-setup, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
```

```{r 21-create-demo-dat}
race <- read_csv("data/21_race.csv", show_col_types = FALSE)
ranking <- read_csv("data/21_ultra_rankings.csv", show_col_types = FALSE)

best_results <- ranking %>%
  filter(rank <= 10) %>%
  group_by(race_year_id) %>%
  summarise(time_in_seconds = mean(time_in_seconds), top_10 = n()) %>%
  filter(top_10 == 10) %>%
  select(-top_10)

race_top_results <- race %>%
  filter(participation == "solo" || participation == "Solo") %>%
  inner_join(best_results, by = "race_year_id") %>%
  mutate(avg_elevation_gain = elevation_gain / distance, avg_velocity = distance / time_in_seconds * 3600) %>%
  filter(distance > 0)

glimpse(race_top_results)
```

We will work with races with non-0 distance, solo participation and at least 10 participants. For each race, the avg velocity is calculated from the velocity of the top 10 racers.

```{r 21-set-seed}
set.seed(345129)
```

## Tidy method from the {broom} package

- predictable outcome for many different models and statistical tests
  - always a tibble
  - consistent column names
- most useful for analysing / visualizing multiple models/tests
  - easier to combine results (no rownames)
- also used internally by higher level functions in tidymodels packages
- other packages also provide tidy methods for their own data structures
- different models, tests will have different structures based on what makes sense, but use as similar structure as possible

You can get the same outcome from many different input formats.

```{r 21-visualize-simple-lm-model}
race_top_results %>%
  ggplot(aes(avg_elevation_gain, avg_velocity)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  expand_limits(y = 0)
```

As makes sense intuitively, higher elevation gain per mile results in lower velocity.

```{r 21-tidy-fit-workflow}
lm_spec <- linear_reg() %>% set_engine("lm")

wf <- workflow() %>%
  add_model(lm_spec) %>%
  add_formula(avg_velocity ~ avg_elevation_gain + distance)

fitted_wf <- wf %>%
  fit(race_top_results)

fitted_wf %>% tidy()
```

```{r 21-tidy-parsnip-model-fit}
lm_spec %>%
  fit(avg_velocity ~ avg_elevation_gain + distance, data = race_top_results) %>%
  tidy()
```

{broom} existed before {tidymodels}, it works for base R lm model object as well.

```{r 21-tidy-bare-lm}
fitted_wf %>% extract_fit_engine() %>% tidy()

lm(avg_velocity ~ avg_elevation_gain + distance, data = race_top_results) %>%
  tidy()
```

In addition to models, we can tidy the result of tests such as correlation test or t-test.

```{r 21-tidy-tests}
cor.test(race_top_results$avg_velocity, race_top_results$avg_elevation_gain) %>%
  tidy()

t.test(
  race_top_results %>% filter(date > '2015-01-01') %>% pull(avg_velocity),
  race_top_results %>% filter(date <= '2015-01-01') %>% pull(avg_velocity)
) %>%
  tidy()
```

## {infer} for simple, high level hypothesis testing

- specify relationship and optionally hypothesis
- calculate statistics from simulation or based on theoretical distributions
- many common tests are supported for continuous and discreet variables as well

### p value for idependence based on simulation with permutation

```{r 21-infer-independence-p}
observed <- race_top_results %>%
  specify(avg_velocity ~ avg_elevation_gain) %>%
  calculate(stat = "correlation")

observed

permuted <- race_top_results %>%
  specify(avg_velocity ~ avg_elevation_gain) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "correlation")

permuted

permuted %>%
  visualize() +
  shade_p_value(observed, direction = "two_sided")

get_p_value(permuted, observed, direction = "two_sided")
```

### Confidence interval for correlation based on simulation with bootstrapping

```{r 21-infer-correlation-ci}
bootstrapped <- race_top_results %>%
  specify(avg_velocity ~ avg_elevation_gain) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "correlation")

bootstrapped %>%
  visualize() +
  shade_confidence_interval(get_confidence_interval(bootstrapped))
```

### Use theory instead of simulation

```{r 21-infer-t-mu}
observed_t <- race_top_results %>%
  specify(response = avg_velocity) %>%
  hypothesise(null = "point", mu = 7) %>%
  calculate(stat = "t")

race_top_results %>%
  specify(response = avg_velocity) %>%
  assume("t") %>%
  visualize() +
  shade_p_value(observed_t, direction = "two_sided")

race_top_results %>%
  specify(response = avg_velocity) %>%
  assume("t") %>%
  get_p_value(observed_t, "two_sided")
```

### Linear models with multiple explanatory variables

```{r 21-infer-fit}
my_formula <- as.formula(avg_velocity ~ aid_stations + participants)

observed_fit <- race_top_results %>%
  specify(my_formula) %>%
  fit()

observed_fit
```


```{r 21-infer-fit-simulate, eval=FALSE}
permuted_fits <- race_top_results %>%
  specify(my_formula) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 1000, type = "permute", variables = c(aid_stations, participants)) %>%
  fit()

bootstrapped_fits <- race_top_results %>%
  specify(my_formula) %>%
  generate(reps = 2000, type = "bootstrap") %>%
  fit()
```

```{r 21-infer-fit-read, echo=FALSE}
permuted_fits <- read_rds("data/21_permuted_fits.rds")
bootstrapped_fits <- read_rds("data/21_bootstrapped_fits.rds")
```

```{r 21-infer-fit-permute-eval}
permuted_fits %>% get_p_value(observed_fit, "two_sided")

visualize(permuted_fits) +
  shade_p_value(observed_fit, "two_sided")
```

```{r 21-infer-fit-bootstrap_eval}
bootstrapped_fits %>%
  get_confidence_interval(type = "percentile", point_estimate = observed_fit)
```

## reg_intervals from {rsample}

Similar purpose (?) as {infer} package, few supported models and interval types

```{r 21-reg-intervals, eval=FALSE}
reg_intervals(
  my_formula,
  race_top_results,
  model_fn = "glm",
  times = 2000,
  type = "percentile",
  keep_reps = TRUE
)
```

```{r 21-reg-intervals-eval, echo=FALSE}
read_rds("data/21_reg_intervals.rds")
```

## Inference with lower level helpers

Most useful for models not supported by anova, or parsnip, etc or when you want to have more control.

```{r 21-infer-lm}
formulas <- list(
  "full" = as.formula(avg_velocity ~ avg_elevation_gain + distance + aid_stations),
  "partial" = as.formula(avg_velocity ~ avg_elevation_gain + distance),
  "minimal" = as.formula(avg_velocity ~ avg_elevation_gain)
)

lm_spec <- linear_reg() %>% set_engine("lm")

AICs <- map(formulas, function(formula) {
  lm_spec %>%
    fit(formula, data = race_top_results) %>%
    extract_fit_engine() %>%
    AIC()
})
AICs
```

How to determine whether these differences are significant? One possible solution is bootstrapping which can also be used when there are no nice theoretical properties.

```{r 21-many-models, eval=FALSE}
velocity_model_summaries <- race_top_results %>%
  bootstraps(times = 1000, apparent = TRUE) %>%
  mutate(
    full_aic = map_dbl(splits, ~ fit(lm_spec, formulas[["full"]], data = analysis(.x)) %>% extract_fit_engine() %>% AIC()),
    partial_aic = map_dbl(splits, ~ fit(lm_spec, formulas[["partial"]], data = analysis(.x)) %>% extract_fit_engine() %>% AIC()),
    minimal_aic = map_dbl(splits, ~ fit(lm_spec, formulas[["minimal"]], data = analysis(.x)) %>% extract_fit_engine() %>% AIC())
  ) %>% 
  select(full_aic, partial_aic, minimal_aic)
```

```{r 21-many-models-read, echo=FALSE}
velocity_model_summaries <- read_rds("data/21_velocity_model_summaries.rds")
```


```{r 21-aic-compare}
velocity_model_summaries %>%
  pivot_longer(c(full_aic, partial_aic, minimal_aic)) %>%
  ggplot(aes(x = name, y = value)) +
  geom_boxplot()

velocity_model_summaries %>%
  summarize(
    full_vs_partial = mean(full_aic < partial_aic),
    partial_vs_minimal = mean(partial_aic < minimal_aic)
  )

```

```{r 21-model-coeffs}
velocity_model_summaries %>%
  unnest(full_coeffs) %>%
  ggplot(aes(x = estimate)) +
  geom_histogram(bins = 5) +
  facet_wrap(~term, scales = "free_x") +
  geom_vline(xintercept = 0, col = "green")
```

Small reps used for speed

```{r 21-model-coef-ci}
race_top_results %>%
  bootstraps(times = 20) %>%
  mutate(
    full_coeffs = map(splits, ~ fit(lm_spec, formulas[["full"]], data = analysis(.x)) %>% tidy())
  ) %>% 
  int_pctl(full_coeffs)
```
The key components are `bootstraps()` and `mutate + map` which give you quite a bit of flexibility to compute many statistics.

## Meeting Videos

### Cohort 3

`r knitr::include_url("https://www.youtube.com/embed/p7CyaalpUNI")`

<details>
  <summary> Meeting chat log </summary>
  
```
00:17:29	Federica Gazzelloni:	tidy function if you google it, this is what it pops up at first: https://www.rdocumentation.org/packages/broom/versions/0.3.4/topics/tidy
00:53:19	Federica Gazzelloni:	calculate function: https://www.rdocumentation.org/packages/infer/versions/0.5.4/topics/calculate
```
</details>

### Cohort 4

`r knitr::include_url("https://www.youtube.com/embed/7BkLFznRko8")`

<details>
  <summary> Meeting chat log </summary>
  
```
00:33:49	Brandon Hurr:	P-value 🤏
00:43:07	Federica Gazzelloni:	moderndive.com
00:43:17	Federica Gazzelloni:	https://infer.tidymodels.org/
00:57:16	Federica Gazzelloni:	https://juliasilge.com/blog/rstats-vignettes/
```
</details>

`r knitr::include_url("https://www.youtube.com/embed/2TgtJXNrk_c")`

<details>
  <summary> Meeting chat log </summary>
  
```
00:23:14	Federica Gazzelloni:	total # of parameters: H(P+1)+H+1
00:24:15	Federica Gazzelloni:	“For this type of network model and P predictors, there are a total of H(P + 1) + H + 1 total parameters being estimated, which quickly becomes large as P increases. (APM pg.143)"
00:25:39	Federica Gazzelloni:	The name neural network originally derived from thinking of these hidden units as analogous to neurons in the brain (ISLR pg 405)
00:46:35	Stephen.Charlesworth:	I'm installing rTools right now :)
```
</details>
